\thetitle

### 2013-04-21 tfpie/tfpie.tex
something submittable

No SE tools/mechanism built into the language.

\subsection{Universe}

\subsection{Programming to interfaces}

>>>>

Our introduction to object-oriented programming is built on a series
of ``language levels'', each of which introduces additional features,
adding complexity to the programming model and expressiveness to the
programs.  Each language is  \texttt{class/}\textit{N}
for some \textit{N}, with features appearing in the following order.
\begin{enumerate}
  \setcounter{enumi}{-1}
\item Classes and Objects
\item Abbreviated notation for method calls
\item Super classes
\item Overriding
\item Constructors
\end{enumerate}

Several commonalities run through all of these languages.  First, they
are all purely functional; we do not introduce imperative I/O or
side-effects until after transitioning to Java in the second half of
the course.  Second, they all are a super set of the
\emph{Intermediate Student} language from \htdp, meaning that they
support higher-order functional programming and lists.

One key principle that we adhere to in the design of the language
levels is that no features of the language are added purely to support
``software engineering'' concerns such as specification mechanisms.
Not only does that language not support declaring types or contracts,
but interfaces are described purely in comments.

This is not to say that interfaces and contracts are optional; in
fact, they are mandatory.  But the focus of the first part of the
course is on the fundamentals of object-orientation.  Teaching the use
of software engineering tools such as type systems, while vital, is a
topic which we defer to the second half of the course when we
transition to Java.

We made this decision after experience in which students were confused
about the relationship between explicit interface specifications, type
systems, and the informal data definitions and contracts which
students are required to write for all methods.  After removing
interfaces from the language and making them purely a specification
construct, this confusion disappeared.

%% Fixme: add these sections back later

%\subsection{Universe}

%\subsection{Programming to interfaces}
